---
title: "CIPHA_vs_Model"
output: pdf_document
---


CIPHA data vs Model data calculations of stem # per hectare, basal area, biomass, height, and dbh

# Key

1 - Setup (import packages and datasets)

2 - CIPHA (calculations from MEN data, for each species by node and MEN year)
  2a - Stem # calculations
  2b - Other parameters function
  2c - MEN00 calculations
  2d - MEN04 calculations
  2e - MEN08 calculations
  2f - MEN12 calculations
  2g - MEN16 calculations
  2h - Overall averages

3 - Model (calculations from model data, for each species by node)
  3a - Parameters function
  3b - BAT calculations
  3c - EDG calculations
  3d - MIN calculations
  3e - YRK calculations

4 - Historical Data (calculations from CFS and USDA data for each species)
  This can be used for comparison of Parkland region species to species overall

# Section 1 - Setup

This section reads in each package used throughout, as well as the necessary datasets. First, each of the MEN datasets are read in to calculate CIPHA averages. Then, each parameter for the different nodes run in the model are read in.
```{r sec1}
library("readxl")
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE) 
library(ggplot2)
#library(tidyverse)
#iris %>% summarise(n=n())
#iris %>% group_by(Species) %>% summarise(n=n())
#iris %>% group_by(Species) %>% summarise(n=n(), .groups = "keep")
#iris %>% group_by(Species) %>% summarise(n=n(), .groups = "drop")
#iris %>% group_by(Species) %>% summarise(n=n(), .groups = "drop_last")
#iris %>% group_by(Species) %>% summarise(n=n(), .groups = "rowwise")

#import MEN datasets
setwd("/Users/alexcreighton/Desktop/NASA/codes/CIPHA/MEN_Parkland")
MEN00<-read_excel("MEN00_Parkland.xlsx")
MEN04<-read_excel("MEN04_Parkland.xlsx")
MEN08<-read_excel("MEN08_Parkland.xlsx")
MEN12<-read_excel("MEN12_Parkland.xlsx")
MEN16<-read_excel("MEN16_Parkland.xlsx")

#import model datasets
###BAT
setwd("/Users/alexcreighton/Desktop/NASA/runs/BAT12/BAT12_2020-08-28_mf10")
BAT_stems<-read.csv("stems.csv")
BAT_basal<-read.csv("basal.csv")
BAT_bio<-read.csv("bio.csv")
#BAT_height<-read.csv("height.csv")
BAT_dbh<-read.csv("dbh.csv")

###EDG
setwd("/Users/alexcreighton/Desktop/NASA/runs/EDG22/EDG22_2020-08-28_mf10")
EDG_stems<-read.csv("stems.csv")
EDG_basal<-read.csv("basal.csv")
EDG_bio<-read.csv("bio.csv")
#EDG_height<-read.csv("height.csv")
EDG_dbh<-read.csv("dbh.csv")

###MIN
setwd("/Users/alexcreighton/Desktop/NASA/runs/MIN22/MIN22_2020-08-28_mf10")
MIN_stems<-read.csv("stems.csv")
MIN_basal<-read.csv("basal.csv")
MIN_bio<-read.csv("bio.csv")
#MIN_height<-read.csv("height.csv")
MIN_dbh<-read.csv("dbh.csv")

###YRK (1-1 ONLY)
setwd("/Users/alexcreighton/Desktop/NASA/runs/YRK11/YRK11_2020-11-16_mf10")
YRK_stems<-read.csv("stems.csv")
YRK_basal<-read.csv("basal.csv")
YRK_bio<-read.csv("bio.csv")
#YRK_height<-read.csv("height.csv")
YRK_dbh<-read.csv("dbh.csv")

```



# Section 2 - CIPHA Calculations

This section contains the calculations for each of the parameters being analyzed from the MEN datasets. The calculations can be done using one function created with the intention of inputting the desired dataset, node, grouping variable, and parameter and providing a table with the calculations. This section is broken up into subsections - the first calculating stem #, followed by the function for other parameters, then each MEN year's results.

###2a - Stem # per Hectare
```{r sec2a}
species_count <- function(df, desired_node, grouping_var) {
  grouping_var = enquo(grouping_var)
  
  count <- df %>%
    filter(node==desired_node)%>%
    group_by(!!grouping_var)%>%
    count(!!grouping_var)
}

#MEN00
test<-species_count(MEN00, "BAT", species)
View(test)
#304 - 4, 311 - 209
test<-species_count(MEN00, "EDG", species)
View(test)
#304 - 1, 311 - 200
test<-species_count(MEN00, "MIN", species)
View(test)
#168 - 10, 304 - 20, 311 - 235
test<-species_count(MEN00, "YRK", species)
View(test)
#304 - 3, 311 - 167

#MEN04
test<-species_count(MEN04, "BAT", species)
View(test)
#304 - 4, 311 - 209
test<-species_count(MEN04, "EDG", species)
View(test)
#304 - 1, 311 - 200
test<-species_count(MEN04, "MIN", species)
View(test)
#168 - 10, 304 - 20, 311 - 235
test<-species_count(MEN04, "YRK", species)
View(test)
#304 - 4, 311 - 170

#MEN08
test<-species_count(MEN08, "BAT", species)
View(test)
#304 - 2, 311 - 148
test<-species_count(MEN08, "EDG", species)
View(test)
#304 - 1, 311 - 200
test<-species_count(MEN08, "MIN", species)
View(test)
#168 - 10, 304 - 19, 311 - 236
test<-species_count(MEN08, "YRK", species)
View(test)
#304 - 4, 311 - 170

#MEN12
test<-species_count(MEN12, "BAT", species)
View(test)
#304 - 4, 311 - 210
test<-species_count(MEN12, "EDG", species)
View(test)
#304 - 1, 311 - 209
test<-species_count(MEN12, "MIN", species)
View(test)
#168 - 15, 304 - 22, 311 - 258
test<-species_count(MEN12, "YRK", species)
View(test)
#304 - 4, 311 - 176

#MEN16
test<-species_count(MEN16, "BAT", species)
View(test)
#304 - 4, 311 - 213
test<-species_count(MEN16, "EDG", species)
View(test)
#304 - 1, 311 - 223
test<-species_count(MEN16, "MIN", species)
View(test)
#168 - 17, 304 - 22, 311 - 276
test<-species_count(MEN16, "YRK", species)
View(test)
#304 - 4, 311 - 176

```


###2b - Other parameters' function
```{r sec2b}
#calculate biomass for each dataset and species within the dataset
MEN00$biomass[MEN00$species==168]<-0.1182*(MEN00$dbh)^2.4287
MEN00$biomass[MEN00$species==304]<-0.0510*(MEN00$dbh)^2.4529
MEN00$biomass[MEN00$species==311]<-0.1008*(MEN00$dbh)^2.4341

MEN04$biomass[MEN04$species==168]<-0.1182*(MEN04$dbh)^2.4287
MEN04$biomass[MEN04$species==304]<-0.0510*(MEN04$dbh)^2.4529
MEN04$biomass[MEN04$species==311]<-0.1008*(MEN04$dbh)^2.4341

MEN08$biomass[MEN08$species==168]<-0.1182*(MEN08$dbh)^2.4287
MEN08$biomass[MEN08$species==304]<-0.0510*(MEN08$dbh)^2.4529
MEN08$biomass[MEN08$species==311]<-0.1008*(MEN08$dbh)^2.4341

MEN12$biomass[MEN12$species==168]<-0.1182*(MEN12$dbh)^2.4287
MEN12$biomass[MEN12$species==304]<-0.0510*(MEN12$dbh)^2.4529
MEN12$biomass[MEN12$species==311]<-0.1008*(MEN12$dbh)^2.4341

MEN16$biomass[MEN16$species==168]<-0.1182*(MEN16$dbh)^2.4287
MEN16$biomass[MEN16$species==304]<-0.0510*(MEN16$dbh)^2.4529
MEN16$biomass[MEN16$species==311]<-0.1008*(MEN16$dbh)^2.4341

#calculate basal area for each dataset
MEN00$basal<-0.005454*MEN00$dbh^2
MEN04$basal<-0.005454*MEN04$dbh^2
MEN08$basal<-0.005454*MEN08$dbh^2
MEN12$basal<-0.005454*MEN12$dbh^2
MEN16$basal<-0.005454*MEN16$dbh^2


species_avgs_cipha <- function(df, desired_node, grouping_var, target_var) {
  target_var = enquo(target_var)
  grouping_var = enquo(grouping_var)
  
  averages <- df %>%
    filter(node==desired_node)%>%
    group_by(!!grouping_var)%>%
    summarise(var_avg=mean(!!target_var,na.rm = TRUE))
}


```


###2c - MEN00
```{r sec2c}
##MEN00
###height
test<-species_avgs_cipha(MEN00, "BAT", species, ht_total)
View(test)
#304 - 12.6; 311 - 13.7866

test<-species_avgs_cipha(MEN00, "EDG", species, ht_total)
View(test)
#311 - 13.552

test<-species_avgs_cipha(MEN00, "MIN", species, ht_total)
View(test)
#168 - 3.67; 304 - 6.52; 311 - 14.89786

test<-species_avgs_cipha(MEN00, "YRK", species, ht_total)
View(test)
#304 - 11.3; 311 - 14.96108


###dbh
test<-species_avgs_cipha(MEN00, "BAT", species, dbh)
View(test)
#304 - 23.25; 311 - 18.80861

test<-species_avgs_cipha(MEN00, "EDG", species, dbh)
View(test)
#304 - 10.0; 311 - 17.475

test<-species_avgs_cipha(MEN00, "MIN", species, dbh)
View(test)
#168 - 9.10; 304 - 14.10; 311 - 18.17447

test<-species_avgs_cipha(MEN00, "YRK", species, dbh)
View(test)
#304 - 14.3333; 311 - 16.13174


###basal area
test<-species_avgs_cipha(MEN00, "BAT", species, basal)
View(test)
#304 - 3.041968; 311 - 2.061847

test<-species_avgs_cipha(MEN00, "EDG", species, basal)
View(test)
#304 - 0.545400; 311 - 1.845199

test<-species_avgs_cipha(MEN00, "MIN", species, basal)
View(test)
#168 - 0.5197662; 304 - 1.1911536; 311 - 1.9773883

test<-species_avgs_cipha(MEN00, "YRK", species, basal)
View(test)
#304 - 1.20897; 311 - 1.55488


###biomass
test<-species_avgs_cipha(MEN00, "BAT", species, biomass)
View(test)
#304 - 95.9746; 311 - 142.9867

test<-species_avgs_cipha(MEN00, "EDG", species, biomass)
View(test)
#304 - 89.29803; 311 - 128.18823

test<-species_avgs_cipha(MEN00, "MIN", species, biomass)
View(test)
#168 - 175.38714; 304 - 88.15247; 311 - 124.28361

test<-species_avgs_cipha(MEN00, "YRK", species, biomass)
View(test)
#304 - 65.83038; 311 - 119.60756

```


###2d - MEN04
```{r sec2c}
##MEN04
###height
test<-species_avgs_cipha(MEN04, "BAT", species, ht_total)
View(test)
#304 - 15.86667; 311 - 15.05714

test<-species_avgs_cipha(MEN04, "EDG", species, ht_total)
View(test)
#311 - 14.79939

test<-species_avgs_cipha(MEN04, "MIN", species, ht_total)
View(test)
#168 - 4.000000; 304 - 6.533333; 311 - 16.702273

test<-species_avgs_cipha(MEN04, "YRK", species, ht_total)
View(test)
#304 - 10.775; 311 - 15.27607


###dbh
test<-species_avgs_cipha(MEN04, "BAT", species, dbh)
View(test)
#304 - 21.83333; 311 - 20.48061

test<-species_avgs_cipha(MEN04, "EDG", species, dbh)
View(test)
#311 - 18.71988

test<-species_avgs_cipha(MEN04, "MIN", species, dbh)
View(test)
#168 - 9.666667; 304 - 15.236364; 311 - 19.210048

test<-species_avgs_cipha(MEN04, "YRK", species, dbh)
View(test)
#304 - 12.82500; 311 - 16.51647


###basal area
test<-species_avgs_cipha(MEN04, "BAT", species, basal)
View(test)
#304 - 2.611102; 311 - 2.410661

test<-species_avgs_cipha(MEN04, "EDG", species, basal)
View(test)
#311 - 2.094679

test<-species_avgs_cipha(MEN04, "MIN", species, basal)
View(test)
#168 - 0.549436; 304 - 1.344342; 311 - 2.158886

test<-species_avgs_cipha(MEN04, "YRK", species, basal)
View(test)
#304 - 1.018030; 311 - 1.631481


###biomass
test<-species_avgs_cipha(MEN04, "BAT", species, biomass)
View(test)
#304 - 102.2195; 311 - 170.6453

test<-species_avgs_cipha(MEN04, "EDG", species, biomass)
View(test)
#304 - 87.22636; 311 - 148.93591

test<-species_avgs_cipha(MEN04, "MIN", species, biomass)
View(test)
#168 - 188.3509; 304 - 104.6622; 311 - 140.6343

test<-species_avgs_cipha(MEN04, "YRK", species, biomass)
View(test)
#304 - 75.76029; 311 - 130.25612

```


###2e - MEN08
```{r sec2c}
##MEN08
###height - NO DATA
test<-species_avgs_cipha(MEN08, "BAT", species, ht_total)
View(test)

test<-species_avgs_cipha(MEN08, "EDG", species, ht_total)
View(test)

test<-species_avgs_cipha(MEN08, "MIN", species, ht_total)
View(test)

test<-species_avgs_cipha(MEN08, "YRK", species, ht_total)
View(test)


###dbh
test<-species_avgs_cipha(MEN08, "BAT", species, dbh)
View(test)
#304 - 27.1; 311 - 20.53393

test<-species_avgs_cipha(MEN08, "EDG", species, dbh)
View(test)
#311 - 19.32704

test<-species_avgs_cipha(MEN08, "MIN", species, dbh)
View(test)
#168 - 10.00000; 304 - 15.40000; 311 - 20.60171

test<-species_avgs_cipha(MEN08, "YRK", species, dbh)
View(test)
#304 - 13.45000; 311 - 17.61437


###basal area
test<-species_avgs_cipha(MEN08, "BAT", species, basal)
View(test)
#304 - 4.0648662; 311 - 2.4847051

test<-species_avgs_cipha(MEN08, "EDG", species, basal)
View(test)
#311 - 2.240644

test<-species_avgs_cipha(MEN08, "MIN", species, basal)
View(test)
#168 - 0.5487088; 304 - 1.3935437; 311 - 2.4541033

test<-species_avgs_cipha(MEN08, "YRK", species, basal)
View(test)
#304 - 1.096963; 311 - 1.836774


###biomass
test<-species_avgs_cipha(MEN08, "BAT", species, biomass)
View(test)
#304 - 37.22913; 311 - 182.28589

test<-species_avgs_cipha(MEN08, "EDG", species, biomass)
View(test)
#304 - 62.01974; 311 - 159.60239

test<-species_avgs_cipha(MEN08, "MIN", species, biomass)
View(test)
#168 - 189.1799; 304 - 104.9069; 311 - 168.6125

test<-species_avgs_cipha(MEN08, "YRK", species, biomass)
View(test)
#304 - 118.6498; 311 - 141.7703

```


###2f - MEN12
```{r sec2c}
##MEN12
###height
test<-species_avgs_cipha(MEN12, "BAT", species, ht_total)
View(test)
#304 - 5.400000; 311 - 4.009615

test<-species_avgs_cipha(MEN12, "EDG", species, ht_total)
View(test)
#311 - 11.5043

test<-species_avgs_cipha(MEN12, "MIN", species, ht_total)
View(test)
#168 - 1.900000; 304 - 2.650000; 311 - 4.138095

test<-species_avgs_cipha(MEN12, "YRK", species, ht_total)
View(test)
#304 - 3.900000; 311 - 5.142857


###dbh
test<-species_avgs_cipha(MEN12, "BAT", species, dbh)
View(test)
#304 - 22.20000; 311 - 22.32464

test<-species_avgs_cipha(MEN12, "EDG", species, dbh)
View(test)
#311 - 20.50347

test<-species_avgs_cipha(MEN12, "MIN", species, dbh)
View(test)
#168 - 9.30000; 304 - 13.84286; 311 - 19.97738

test<-species_avgs_cipha(MEN12, "YRK", species, dbh)
View(test)
#304 - 14.15; 311 - 17.84658


###basal area
test<-species_avgs_cipha(MEN12, "BAT", species, basal)
View(test)
#304 - 2.7070384; 311 - 2.8940212

test<-species_avgs_cipha(MEN12, "EDG", species, basal)
View(test)
#311 - 2.541984

test<-species_avgs_cipha(MEN12, "MIN", species, basal)
View(test)
#168 - 0.4796793; 304 - 1.2423199; 311 - 2.4306030

test<-species_avgs_cipha(MEN12, "YRK", species, basal)
View(test)
#304 - 1.197508; 311 - 1.901423


###biomass
test<-species_avgs_cipha(MEN12, "BAT", species, biomass)
View(test)
#304 - 131.0696; 311 - 211.5177

test<-species_avgs_cipha(MEN12, "EDG", species, biomass)
View(test)
#304 - 81.18327; 311 - 189.04279

test<-species_avgs_cipha(MEN12, "MIN", species, biomass)
View(test)
#168 - 245.5263; 304 - 137.4687; 311 - 159.5103

test<-species_avgs_cipha(MEN12, "YRK", species, biomass)
View(test)
#304 - 67.81865; 311 - 158.02847

```


###2g - MEN16
```{r sec2c}
##MEN16
###height - NO DATA
test<-species_avgs_cipha(MEN16, "BAT", species, ht_total)
View(test)

test<-species_avgs_cipha(MEN16, "EDG", species, ht_total)
View(test)

test<-species_avgs_cipha(MEN16, "MIN", species, ht_total)
View(test)

test<-species_avgs_cipha(MEN16, "YRK", species, ht_total)
View(test)


###dbh
test<-species_avgs_cipha(MEN16, "BAT", species, dbh)
View(test)
#304 - 22.5000; 311 - 24.1266

test<-species_avgs_cipha(MEN16, "EDG", species, dbh)
View(test)
#311 - 20.35455

test<-species_avgs_cipha(MEN16, "MIN", species, dbh)
View(test)
#168 - 9.72500; 304 - 14.24286; 311 - 19.80888

test<-species_avgs_cipha(MEN16, "YRK", species, dbh)
View(test)
#311 - 18.54015


###basal area
test<-species_avgs_cipha(MEN16, "BAT", species, basal)
View(test)
#304 - 2.797957; 311 - 3.399271

test<-species_avgs_cipha(MEN16, "EDG", species, basal)
View(test)
#311 - 2.600805

test<-species_avgs_cipha(MEN16, "MIN", species, basal)
View(test)
#168 - 0.5270609; 304 - 1.3180526; 311 - 2.4624787

test<-species_avgs_cipha(MEN16, "YRK", species, basal)
View(test)
#311 - 2.041534


###biomass
test<-species_avgs_cipha(MEN16, "BAT", species, biomass)
View(test)
#304 - 200.4508; 311 - 262.5749

test<-species_avgs_cipha(MEN16, "EDG", species, biomass)
View(test)
#311 - 200.984

test<-species_avgs_cipha(MEN16, "MIN", species, biomass)
View(test)
#168 - 487.5925; 304 - 194.2675; 311 - 158.2773

test<-species_avgs_cipha(MEN16, "YRK", species, biomass)
View(test)
#311 - 177.0033

```



# Section 3 - Model Calculations

This section contains the calculations for each of the parameters being analyzed from the model's output datasets. This section is broken up into subsections - the first is the overall function to calculation parameter averages, followed by each nodes' results.

### 3a - Parameter function
```{r sec3a}
species_avgs_model <- function(df, target_species) {
  target_species = enquo(target_species)
  
  averages <- df %>%
    summarise(var_avg=mean(!!target_species,na.rm = TRUE))
}

```


### 3b - BAT
```{r sec3b}
##basal
test<-species_avgs_model(BAT_basal, BEPA)
View(test)
test<-species_avgs_model(BAT_basal, POBA)
View(test)
test<-species_avgs_model(BAT_basal, POTR)
View(test)


##biomass
test<-species_avgs_model(BAT_bio, BEPA)
View(test)
test<-species_avgs_model(BAT_bio, POBA)
View(test)
test<-species_avgs_model(BAT_bio, POTR)
View(test)


##DBH
test<-species_avgs_model(BAT_dbh, BEPA)
View(test)
test<-species_avgs_model(BAT_dbh, POBA)
View(test)
test<-species_avgs_model(BAT_dbh, POTR)
View(test)


##stems
test<-species_avgs_model(BAT_stems, BEPA)
View(test)
test<-species_avgs_model(BAT_stems, POBA)
View(test)
test<-species_avgs_model(BAT_stems, POTR)
View(test)

```


### 3c - EDG
```{r sec3c}
##basal
test<-species_avgs_model(EDG_basal, BEPA)
View(test)
test<-species_avgs_model(EDG_basal, POBA)
View(test)
test<-species_avgs_model(EDG_basal, POTR)
View(test)


##biomass
test<-species_avgs_model(EDG_bio, BEPA)
View(test)
test<-species_avgs_model(EDG_bio, POBA)
View(test)
test<-species_avgs_model(EDG_bio, POTR)
View(test)


##DBH
test<-species_avgs_model(EDG_dbh, BEPA)
View(test)
test<-species_avgs_model(EDG_dbh, POBA)
View(test)
test<-species_avgs_model(EDG_dbh, POTR)
View(test)


##stems
test<-species_avgs_model(EDG_stems, BEPA)
View(test)
test<-species_avgs_model(EDG_stems, POBA)
View(test)
test<-species_avgs_model(EDG_stems, POTR)
View(test)

```


### 3d - MIN
```{r sec3d}
##basal
test<-species_avgs_model(MIN_basal, BEPA)
View(test)
test<-species_avgs_model(MIN_basal, POBA)
View(test)
test<-species_avgs_model(MIN_basal, POTR)
View(test)


##biomass
test<-species_avgs_model(MIN_bio, BEPA)
View(test)
test<-species_avgs_model(MIN_bio, POBA)
View(test)
test<-species_avgs_model(MIN_bio, POTR)
View(test)


##DBH
test<-species_avgs_model(MIN_dbh, BEPA)
View(test)
test<-species_avgs_model(MIN_dbh, POBA)
View(test)
test<-species_avgs_model(MIN_dbh, POTR)
View(test)


##stems
test<-species_avgs_model(MIN_stems, BEPA)
View(test)
test<-species_avgs_model(MIN_stems, POBA)
View(test)
test<-species_avgs_model(MIN_stems, POTR)
View(test)

```


### 3e - YRK
```{r sec3e}
##basal
test<-species_avgs_model(YRK_basal, BEPA)
View(test)
test<-species_avgs_model(YRK_basal, POBA)
View(test)
test<-species_avgs_model(YRK_basal, POTR)
View(test)


##biomass
test<-species_avgs_model(YRK_bio, BEPA)
View(test)
test<-species_avgs_model(YRK_bio, POBA)
View(test)
test<-species_avgs_model(YRK_bio, POTR)
View(test)


##DBH
test<-species_avgs_model(YRK_dbh, BEPA)
View(test)
test<-species_avgs_model(YRK_dbh, POBA)
View(test)
test<-species_avgs_model(YRK_dbh, POTR)
View(test)


##stems
test<-species_avgs_model(YRK_stems, BEPA)
View(test)
test<-species_avgs_model(YRK_stems, POBA)
View(test)
test<-species_avgs_model(YRK_stems, POTR)
View(test)

```




# Section 4

This section contains calculations from historical data (from the CFS and USDA) for the relevant species in the Parkland region. This information can be used for comparisons of Parkland species to the species' overall. 

```{r sec4}
#not ultimately used

```

